---
title: "RXFP3 Antagonist RNA-seq: focus on hypothalamus: prep counts"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 7
theme: cosmo
---

Source: https://github.com/markziemann/rxfp3_ant

# Introduction

This script is just for preparing the counts into a useful format.

## Bioinformatics methods

Fastqc (v0.11.9) was used to inspect sequence quality[1].

The mouse transcriptome was downloaded from GENCODE version 28[2].

Skewer (v0.2.2) was used to trim low quality bases (qual<20) from the 3' end of the read[3].

Kallisto (0.46.1) was used to map RNA-seq reads to the transcriptome [4].

Multiqc was used to tabulate sequence quality, trimming and mapping statistics [5].

Data were read into R v4.1.2 and duplicate lane data were aggregated, and transcript level counts were aggregated to gene level counts.

Genes with an average of less than 10 reads across samples were excluded from downstream analysis.

DESeq (1.32.0) was used with default settings to assay differential expression between control and treatment groups for all tissues [6].

Pathway analysis was performed with reactome gene sets obtained from MSigDB and converted to mouse gene identifiers with the msigdbr package 
(performed on 16-02-2022) [7,8,9].

Differential pathway analysis was performed with the "mitch" bioconductor package [10].

Genes and pathways with false discovery rate (FDR)<0.05 were considered significant.

```{r,packages}

suppressPackageStartupMessages({
    library("zoo")
    library("tidyverse")
    library("reshape2")
    library("DESeq2")
    library("gplots")
    library("fgsea")
    library("MASS")
    library("mitch")
    library("eulerr")
    library("limma")
    library("topconfects")
    library("kableExtra")
    library("vioplot")
    library("beeswarm")
})

```

## Import read counts

Importing RNA-seq data

```{r,importdata}

tmp <- read.table("hypothalamus_3col.tsv.gz",header=F)
x <- as.matrix(acast(tmp, V2~V1, value.var="V3", fun.aggregate = sum))
x <- as.data.frame(x)
accession <- sapply((strsplit(rownames(x),"\\|")),"[[",2)
symbol<-sapply((strsplit(rownames(x),"\\|")),"[[",6)
x$geneid <- paste(accession,symbol)
xx <- aggregate(. ~ geneid,x,sum)
rownames(xx) <- xx$geneid
xx$geneid = NULL
xx <- round(xx)
xx <- xx[,which(colnames(xx)!="test")]
xx[1:6,1:6]
dim(xx)

```

Fix the sample names.

They are duplicated for lane 1 and 2, which I will aggregate.

```{r,colnames}

txx <- as.data.frame(t(xx))
txx$label <- sapply(strsplit(rownames(txx),"_"),"[[",1)
txx[1:3,c(1:4,ncol(txx))]
txx2 <- aggregate(. ~ label,txx,sum)
txx2[1:3,1:4]
rownames(txx2) <- txx2[,1]
txx2[,1] = NULL
xx <- as.data.frame(t(txx2))
xx[1:4,1:5]
write.table(xx,file="hypothalamus_counts.tsv",sep="\t",quote=FALSE)
rxx <- xx/colSums(xx) *1e6
rxx[1:4,1:5]

```

## Session information

```{r,sessioninfo}

sessionInfo()

```
